Parametricity for Primitive Nested Types and GADTs

### 2021-07-14 21-lmcs/21-lmcs.tex
MFPS revisions final

\author[E.~Ghiorzi]{Enrico Ghiorzi}	
\address{Appalachian State University}	
\email{ghiorzie@appstate.edu}

@@

      \alpha})_\emptyset\, \mathit{swap})_\alpha\, t')}
:  \Nat^{\alpha} (\alpha + \mathit{PTree}\,(\alpha \times

@@

    F}\rho'}$ is defined by $\setsem{\Gamma;\Phi \vdash \phi

@@

      F}\rho}}$.  The latter equality holds because $\rho\phi$ and

@@

{\color{red} CHECK CONDITIONS} Using $T_H$, we can define the
functorial action of relational interpretation.

@@

\item The application case is proved by the following sequence of
  equalities, where the second equality is by the induction hypothesis
  and the definition of the relation environment $\Eq_\rho$, the third
  is by the definition of application of relation transformers from
  Definition~\ref{def:rel-transf}, and the fourth is by
  Equation~\ref{eq:eq-pres-eq}:
\[
\begin{split}
\relsem{\Gamma; \Phi \vdash \phi\ol{F}}\Eq_{\rho} &=
(\Eq_{\rho}\phi)\ol{\relsem{\Gamma; \Phi \vdash F}
\Eq_{\rho}}\\
&= \Eq_{\rho \phi}\, \ol{\Eq_{\setsem{\Gamma; \Phi \vdash F}
  \rho}}\\
&= (\Eq_{\rho \phi})^* \,\ol{\Eq_{\setsem{\Gamma; \Phi \vdash F}
  \rho}}\\
&= \Eq_{(\rho \phi) \,\ol{\setsem{\Gamma; \Phi \vdash F} \rho}}\\
&= \Eq_{\setsem{\Gamma; \Phi \vdash \phi\ol{F}}\rho}
\end{split}
\]
\item  The fixpoint case is proven by the sequence of equalities
\[
\begin{split}
\relsem{\Gamma; \Phi \vdash (\mu \phi.\lambda
  \ol{\alpha}. H)\ol{F}}\Eq_{\rho} 
&=(\mu {T_{H, \Eq_{\rho}}}) \,\ol{\relsem{\Gamma; \Phi \vdash F}\Eq_{\rho}}\\ 
&= \colim{n \in \nat}{T^n_{H, \Eq_{\rho}} K_0}\, \ol{\relsem{\Gamma; \Phi 
  \vdash F}\Eq_{\rho}}\\
&= \colim{n \in \nat}{ T^n_{H, \Eq_{\rho}} K_0 \,\ol{\Eq_{\setsem{\Gamma;
    \Phi \vdash F}\rho}}}\\
&= \colim{n \in \nat}{(\Eq_{(T^\set_{H,\rho})^n K_0})^*
  \ol{\Eq_{\setsem{\Gamma; \Phi \vdash F}\rho}}}\\
&= \colim{n \in \nat}{\Eq_{(T^\set_{H,\rho})^n K_0 \,\ol{\setsem{\Gamma;
        \Phi \vdash F}\rho}}}\\ 
&= \Eq_{\colim{n \in \nat}{ (T^\set_{H,\rho})^n K_0\,
    \ol{\setsem{\Gamma; \Phi \vdash F}\rho}}}\\
&= \Eq_{\setsem{\Gamma; \Phi \vdash (\mu \phi.\lambda
      \ol{\alpha}. H)\ol{F}}\rho}
\end{split}
\]
Here, the third equality is by induction hypothesis, the fifth is by
Equation~\ref{eq:eq-pres-eq}, and the fourth equality is because, for
every $n \in \nat$, the following two statements can be proved by
simultaneous induction: for any $H$, $\rho$, and $A$, 
\begin{equation}\label{eq:iel-fix-point-intermediate1}
T^n_{H, \Eq_{\rho}} K_0\, \ol{\Eq_A} = (\Eq_{(T^\set_{H, \rho})^n K_0})^*
\ol{\Eq_A}
\end{equation}
and for any subformula $J$ of $H$,
\begin{equation}\label{eq:iel-fix-point-intermediate2}
\begin{split}
& \relsem{\Gamma; \phi, \ol{\alpha} \vdash J} \Eq_{\rho} [\phi
    := T^{n}_{H,\Eq_{\rho}} K_0] \overline{[\alpha := \Eq_A]} \\
=\;\; & \relsem{\Gamma; \phi, \ol{\alpha} \vdash J} \Eq_{\rho}
[\phi := \Eq_{(T^\set_{H,\rho})^n K_0}] \overline{[\alpha := \Eq_A]}
\end{split}
\end{equation}
We prove Equation~\ref{eq:iel-fix-point-intermediate1} by induction on
$n$.  The case $n=0$ is trivial, because $T^0_{H,\Eq_{\rho}} K_0 =
K_0$ and $(T^\set_{H,\rho})^0 K_0 = K_0$; the inductive step is proved
by the following sequence of equalities:
\[
\begin{split}
T^{n+1}_{H,\Eq_{\rho}} K_0\, \overline{\Eq_A}
&= T^\rel_{H,\Eq_{\rho}} (T^{n}_{H,\Eq_{\rho}} K_0)
\overline{\Eq_A} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash H} \Eq_{\rho} [\phi
  := T^{n}_{H, \Eq_{\rho}} K_0] \overline{[\alpha :=
    \Eq_A]} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash H} \Eq_{\rho} [\phi
  := \Eq_{(T^\set_{H,\rho})^n K_0}] \overline{[\alpha :=
    \Eq_A]} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash H} \Eq_{\rho [\phi
    := (T^\set_{H,\rho})^n K_0] \overline{[\alpha :=
      A]}} \\ 
&= \Eq_{\setsem{\Gamma; \phi, \ol{\alpha} \vdash H} \rho [\phi
    := (T^\set_{H,\rho})^n K_0] \overline{[\alpha :=
      A]}} \\ 
&= \Eq_{(T^\set_{H,\rho})^{n+1} K_0 \overline{A}} \\ 
&= (\Eq_{(T^\set_{H,\rho})^{n+1} K_0})^*\, \overline{\Eq_A} 
\end{split}
\]
Here, the third equality is by
Equation~\ref{eq:iel-fix-point-intermediate2} for $J = H$, the fifth
by the induction hypothesis of the IEL on $H$, and the last is by
Equation~\ref{eq:eq-pres-eq}.

We prove Equation~\ref{eq:iel-fix-point-intermediate2} by structural
induction on $J$. The only interesting cases, though, are when $J =
\phi \ol{G}$ and when $J = (\mu \psi.\lambda \ol\beta. G)\ol K$.
\begin{itemize}
\item The case $J = \phi \ol G$ is proved by the sequence of equalities:
\[
\begin{split}
& \relsem{\Gamma; \phi, \ol{\alpha} \vdash \phi
    \ol{G}} \Eq_{\rho} [\phi := T^{n}_{H,\Eq_{\rho}} K_0]
  \overline{[\alpha := \Eq_A]}
  \\
&= T^{n}_{H,\Eq_{\rho}} K_0\, \overline{\relsem{\Gamma; 
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho} [\phi :=
      T^{n}_{H,\Eq_{\rho}} K_0] \overline{[\alpha :=
        \Eq_A]}} \\ 
&= T^{n}_{H,\Eq_{\rho}} K_0\, \overline{\relsem{\Gamma; 
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho} [\phi :=
      \Eq_{(T^\set_{H,\rho})^{n} K_0}] \overline{[\alpha :=
        \Eq_A]}} \\ 
&= T^{n}_{H,\Eq_{\rho}} K_0\, \overline{\relsem{\Gamma; 
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho [\phi := (T^\set_{H,\rho})^{n}
        K_0] \overline{[\alpha := A]}}} \\ 
&= T^{n}_{H,\Eq_{\rho}} K_0\, \overline{\Eq_{\setsem{\Gamma;
        \phi, \ol{\alpha} \,\vdash G} \rho [\phi :=
        (T^\set_{H,\rho})^{n} K_0] \overline{[\alpha :=
          A]}}} \\ 
&= (\Eq_{(T^\set_{H,\rho})^{n} K_0})^* \,\overline{\Eq_{\setsem{\Gamma;
        \phi, \ol{\alpha} \,\vdash G} \rho [\phi :=
        (T^\set_{H,\rho})^{n} K_0] \overline{[\alpha :=
          A]}}} \\ 
&= (\Eq_{(T^\set_{H,\rho})^{n} K_0})^* \overline{\relsem{\Gamma;
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho} [\phi :=
      \Eq_{(T^\set_{H,\rho})^{n} K_0}] \overline{[\alpha :=
        \Eq_A]}} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash \phi \ol{G}}
  \Eq_{\rho} [\phi := \Eq_{(T^\set_{H,\rho})^{n} K_0}]
  \overline{[\alpha := \Eq_A]} 
\end{split}
\]
Here, the second equality is by the induction hypothesis for
Equation~\ref{eq:iel-fix-point-intermediate2} on the $G$s, the fourth
is by the induction hypothesis for the IEL on the $G$s, and the fifth
is by the induction hypothesis on $n$ for
Equation~\ref{eq:iel-fix-point-intermediate1}.

@@

  \Eq_{\rho} [\phi := T^{n}_{H,\Eq_{\rho}} K_0]

@@

&= (\mu {T_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}})

@@

  [\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]} \\ 
&= \colim{m \in \nat} {T^m_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}}\,K_0

@@

  [\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}) \\ 
&= \colim{m \in \nat} {T^m_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}}\,K_0

@@

  [\phi := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha := \Eq_A}]}) \\  %%
&= \colim{m \in \nat} {T^m_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha := \Eq_A}]}}\,K_0

@@

  [\phi := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha := \Eq_A}]}) \\ 
&= (\mu {T_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha := \Eq_A}]}})

@@

  [\phi := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha := \Eq_A}]} \\ 

@@

  \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha := \Eq_{A}}]

@@

\Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha :=

@@

    \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha :=

@@

& T^{m+1}_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}\,K_0 \\
&= T_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha
        := \Eq_A}]} (T^{m}_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}}

@@

&= T_{G, \Eq_{\rho}[\phi := T^{n}_{H,\Eq_{\rho}} K_0][\overline{\alpha

@@

    \ol\beta \vdash G} \Eq_{\rho} [\phi := T^{n}_{H,\Eq_{\rho}}

@@

        := \Eq_{(T^{\set}_{H,\rho})^n K_0}][\overline{\alpha :=

@@

    \ol\beta \vdash G} \Eq_{\rho} [\phi := \Eq_{(T^{\set}_{H,\rho})^n

@@

      \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^n

@@

&= T_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^n

@@

&= T^{m+1}_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^n

@@

      \Phi \vdash F \times G}\rho}$

@@

\qed

@@

Expanding the second equation gives that if $x \in \setsem{\Gamma;
  \Phi \vdash F}\rho$ then
\[(\setsem{\Gamma; \Phi \vdash F}\id_{\rho}\, x, \setsem{\Gamma;
  \Phi \vdash F} f\, x) \in \relsem{\Gamma; \Phi \vdash F}\graph{f}\]
Then $\setsem{\Gamma; \Phi \vdash F}\id_{\rho}\, x =
\id_{\setsem{\Gamma; \Phi \vdash F}\rho} x = x$, so for any $x \in
\setsem{\Gamma; \Phi \vdash F}\rho$ we have that $(x, \setsem{\Gamma;
  \Phi \vdash F}f\, x) \in \relsem{\Gamma; \Phi \vdash F}\graph{f}$.
Moreover, $x \in \setsem{\Gamma; \Phi \vdash F}\rho$ if and only if
$(x, \setsem{\Gamma; \Phi \vdash F} f\, x) \in \graph{\setsem{\Gamma;
    \Phi \vdash F}f}$ and, if $x \in \setsem{\Gamma; \Phi \vdash
  F}\rho$ then $(x, \setsem{\Gamma; \Phi \vdash F} f\, x) \in
\relsem{\Gamma; \Phi \vdash F} \graph{f}$. Thus, if $(x, \setsem{\Gamma;
  \Phi \vdash F} f\, x) \in \graph{\setsem{\Gamma; \Phi \vdash F}f}$
then $(x, \setsem{\Gamma; \Phi \vdash F} f\, x) \in \relsem{\Gamma;
  \Phi \vdash F} \graph{f}$, i.e., $\graph{\setsem{\Gamma; \Phi \vdash
    F}f} \subseteq \relsem{\Gamma; \Phi \vdash F} \graph{f}$.  \qed

@@

    \alpha.H)\overline \beta \;F)}$$
    is a natural transformation from
  $\setsem{\Gamma;\emptyset \vdash \emptyset}$ to \[\setsem{\Gamma;
    \emptyset \vdash \Nat^\emptyset\;(\Nat^{\ol{\beta}}\,H[\phi
      :=_{\ol{\beta}} F][\ol{\alpha := 
    \phi.\lambda \overline \alpha.H)\overline \beta\,F)}\] we need only
  show that, for all $\rho : \setenv$, the unique $d :
  \setsem{\Gamma;\emptyset \vdash \emptyset} \rho$, and all $\eta :

@@

      :=_{\ol{\beta}} F][\ol{\alpha := 
        \beta}]\,F} \rho$,

@@

by naturality of $\mathit{fold}_{T^\set_{H,\rho}}\eta$.
To see that, for all $\rho : \setenv$, $d \in
\setsem{\Gamma; \emptyset \vdash \emptyset}\rho$, and $\eta :
\setsem{\Gamma; \emptyset \vdash \Nat^{\ol{\beta}}\,H[\phi :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F}
\rho$,

@@

needed for it to be in $\setsem{\Gamma;\emptyset \vdash

@@

additional condition needed for it to be in $\setsem{\Gamma; \emptyset

@@

Now, since $K$ does not contain constants, $\setsem{ \emptyset;
  \ol\alpha \vdash K } (\pi_2 \rho) [\ol{\alpha := 0}] = 0$. Thus, for
any $\ol{S : \set_0}$ and $\ol{f : \setsem{ \emptyset; \ol\alpha
    \vdash K } (\pi_1 \rho)[\ol{\alpha := S}] \to \setsem{ \Gamma;
    \Phi \vdash A } (\pi_1 \rho)} $ we have relations $\ol{(S, 0, 0)}$
and morphisms $\ol{(f, !) : \relsem{ \emptyset; \ol\alpha \vdash K }
  \rho[\ol{\alpha := (S, 0, 0)}] \to \relsem{ \Gamma; \Phi \vdash A }
  \rho}$, where, for any set $X$, we suppress the sub- and
superscripts and write $!$ for the unique morphism $!^0_X$ from $0$ to
$X$. We then define
\[
k : \setsem{ \Gamma; \Phi \vdash (\Lan^{\ol \alpha}_{\ol{K}} F)\ol{A}
} (\pi_1 \rho) \to \pi_1 \,( \relsem{ \Gamma; \Phi \vdash (\Lan^{\ol
    \alpha}_{\ol{K}} F)\ol{A} } \rho )
\]
to be the unique function given by the universal property of the
colimit representation of $k$'s domain. That is, we take $k$ to be the
unique morphism such that, for any $\ol{S : \set_0}$ and $\ol{f :
  \setsem{ \emptyset; \ol\alpha \vdash K } (\pi_1 \rho)[\ol{\alpha :=
      S}] \to \setsem{ \Gamma; \Phi \vdash A } (\pi_1 \rho)}$, we have
$k \circ j_{\ol{S},\, \ol{f}} = \pi_1 \iota_{\ol{(S, 0, 0)},\, \ol{(f,
    !)}}$.

>>>>

\author[E.~Ghiorzi]{Enrico Ghiorzi}	
\address{Appalachian State University}	
\email{enrico.ghiorzi@gmail.com}

@@

      \alpha})_\emptyset\, \mathit{swap})_\alpha\, t')}\\
\hspace*{0.5in}:  \Nat^{\alpha} (\alpha + \mathit{PTree}\,(\alpha \times

@@

    F}\rho'}$ is defined by \[\setsem{\Gamma;\Phi \vdash \phi

@@

      F}\rho}}\]  The latter equality holds because $\rho\phi$ and

@@

Using $T_H$, we can define the functorial action of relational

@@

\item The application case is proved by the following sequence of
  equalities, where the second equality is by the induction hypothesis
  and the definition of the relation environment $\Eq_\rho$, the third
  is by the definition of application of relation transformers from
  Definition~\ref{def:rel-transf}, and the fourth is by
  Equation~\ref{eq:eq-pres-eq}:
\[
\begin{split}
\relsem{\Gamma; \Phi \vdash \phi\ol{F}}\Eq_{\rho} &=
(\Eq_{\rho}\phi)\ol{\relsem{\Gamma; \Phi \vdash F}
\Eq_{\rho}}\\
&= \Eq_{\rho \phi}\, \ol{\Eq_{\setsem{\Gamma; \Phi \vdash F}
  \rho}}\\
&= (\Eq_{\rho \phi})^* \,\ol{\Eq_{\setsem{\Gamma; \Phi \vdash F}
  \rho}}\\
&= \Eq_{(\rho \phi) \,\ol{\setsem{\Gamma; \Phi \vdash F} \rho}}\\
&= \Eq_{\setsem{\Gamma; \Phi \vdash \phi\ol{F}}\rho}
\end{split}
\]
\item  The fixpoint case is proven by the sequence of equalities
\[
\begin{split}
\relsem{\Gamma; \Phi \vdash (\mu \phi.\lambda
  \ol{\alpha}. H)\ol{F}}\Eq_{\rho} 
&=(\mu {T_{H, \Eq_{\rho}}}) \,\ol{\relsem{\Gamma; \Phi \vdash F}\Eq_{\rho}}\\ 
&= \colim{n \in \nat}{T^n_{H, \Eq_{\rho}} K_0}\, \ol{\relsem{\Gamma; \Phi 
  \vdash F}\Eq_{\rho}}\\
&= \colim{n \in \nat}{ T^n_{H, \Eq_{\rho}} K_0 \,\ol{\Eq_{\setsem{\Gamma;
    \Phi \vdash F}\rho}}}\\
&= \colim{n \in \nat}{(\Eq_{(T^\set_{H,\rho})^n K_0})^*
  \ol{\Eq_{\setsem{\Gamma; \Phi \vdash F}\rho}}}\\
&= \colim{n \in \nat}{\Eq_{(T^\set_{H,\rho})^n K_0 \,\ol{\setsem{\Gamma;
        \Phi \vdash F}\rho}}}\\ 
&= \Eq_{\colim{n \in \nat}{ (T^\set_{H,\rho})^n K_0\,
    \ol{\setsem{\Gamma; \Phi \vdash F}\rho}}}\\
&= \Eq_{\setsem{\Gamma; \Phi \vdash (\mu \phi.\lambda
      \ol{\alpha}. H)\ol{F}}\rho}
\end{split}
\]
Here, the third equality is by induction hypothesis, the fifth is by
Equation~\ref{eq:eq-pres-eq}, and the fourth equality is because, for
every $n \in \nat$, the following two statements can be proved by
simultaneous induction: for any $H$, $\rho$, and $A$, 
\begin{equation}\label{eq:iel-fix-point-intermediate1}
T^n_{H, \Eq_{\rho}} K_0\, \ol{\Eq_A} = (\Eq_{(T^\set_{H, \rho})^n K_0})^*
\ol{\Eq_A}
\end{equation}
and for any subformula $J$ of $H$,
\begin{equation}\label{eq:iel-fix-point-intermediate2}
\begin{split}
& \relsem{\Gamma; \phi, \ol{\alpha} \vdash J} \Eq_{\rho} [\phi
    := T^{n}_{H,\Eq_{\rho}} K_0] \overline{[\alpha := \Eq_A]} \\
=\;\; & \relsem{\Gamma; \phi, \ol{\alpha} \vdash J} \Eq_{\rho}
[\phi := \Eq_{(T^\set_{H,\rho})^n K_0}] \overline{[\alpha := \Eq_A]}
\end{split}
\end{equation}
The case $n=0$ is trivial:
Equation~\ref{eq:iel-fix-point-intermediate1} and
Equation~\ref{eq:iel-fix-point-intermediate2} both hold because
$T^0_{H,\Eq_{\rho}} K_0 = K_0$, $(T^\set_{H,\rho})^0 K_0 = K_0$, and
Equation~\ref{eq:eq-pres-eq} holds.

The inductive case is proved as follows.  We prove
Equation~\ref{eq:iel-fix-point-intermediate1} by the following
sequence of equalities:
\[
\begin{split}
T^{n+1}_{H,\Eq_{\rho}} K_0\, \overline{\Eq_A}
&= T^\rel_{H,\Eq_{\rho}} (T^{n}_{H,\Eq_{\rho}} K_0)
\overline{\Eq_A} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash H} \Eq_{\rho} [\phi
  := T^{n}_{H, \Eq_{\rho}} K_0] \overline{[\alpha :=
    \Eq_A]} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash H} \Eq_{\rho} [\phi
  := \Eq_{(T^\set_{H,\rho})^{n} K_0}] \overline{[\alpha :=
    \Eq_A]} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash H} \Eq_{\rho [\phi
    := (T^\set_{H,\rho})^{n} K_0] \overline{[\alpha :=
      A]}} \\ 
&= \Eq_{\setsem{\Gamma; \phi, \ol{\alpha} \vdash H} \rho [\phi
    := (T^\set_{H,\rho})^{n} K_0] \overline{[\alpha :=
      A]}} \\ 
&= \Eq_{(T^\set_{H,\rho})^{n+1} K_0 \overline{A}} \\ 
&= (\Eq_{(T^\set_{H,\rho})^{n+1} K_0})^*\, \overline{\Eq_A} 
\end{split}
\]
Here, the third equality is by the induction hypothesis of
Equation~\ref{eq:iel-fix-point-intermediate2} for $J = H$, the fifth
by the induction hypothesis of the IEL on $H$, and the last is by
Equation~\ref{eq:eq-pres-eq}.

We prove Equation~\ref{eq:iel-fix-point-intermediate2} by structural
induction on $J$. The only interesting cases, though, are when $J =
\phi \ol{G}$ and when $J = (\mu \psi.\lambda \ol\beta. G)\ol K$.
\begin{itemize}
\item The case $J = \phi \ol G$ is proved by the sequence of equalities:
\[
\begin{split}
& \relsem{\Gamma; \phi, \ol{\alpha} \vdash \phi
    \ol{G}} \Eq_{\rho} [\phi := T^{n+1}_{H,\Eq_{\rho}} K_0]
  \overline{[\alpha := \Eq_A]}
  \\
&= T^{n+1}_{H,\Eq_{\rho}} K_0\, \overline{\relsem{\Gamma; 
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho} [\phi :=
      T^{n+1}_{H,\Eq_{\rho}} K_0] \overline{[\alpha :=
        \Eq_A]}} \\ 
&= T^{n+1}_{H,\Eq_{\rho}} K_0\, \overline{\relsem{\Gamma; 
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho} [\phi :=
      \Eq_{(T^\set_{H,\rho})^{n+1} K_0}] \overline{[\alpha :=
        \Eq_A]}} \\ 
&= T^{n+1}_{H,\Eq_{\rho}} K_0\, \overline{\relsem{\Gamma; 
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho [\phi := (T^\set_{H,\rho})^{n+1}
        K_0] \overline{[\alpha := A]}}} \\ 
&= T^{n+1}_{H,\Eq_{\rho}} K_0\, \overline{\Eq_{\setsem{\Gamma;
        \phi, \ol{\alpha} \,\vdash G} \rho [\phi :=
        (T^\set_{H,\rho})^{n+1} K_0] \overline{[\alpha :=
          A]}}} \\ 
&= (\Eq_{(T^\set_{H,\rho})^{n+1} K_0})^* \,\overline{\Eq_{\setsem{\Gamma;
        \phi, \ol{\alpha} \,\vdash G} \rho [\phi :=
        (T^\set_{H,\rho})^{n+1} K_0] \overline{[\alpha :=
          A]}}} \\ 
&= (\Eq_{(T^\set_{H,\rho})^{n+1} K_0})^* \overline{\relsem{\Gamma;
      \phi, \ol{\alpha} \vdash G} \Eq_{\rho} [\phi :=
      \Eq_{(T^\set_{H,\rho})^{n+1} K_0}] \overline{[\alpha :=
        \Eq_A]}} \\ 
&= \relsem{\Gamma; \phi, \ol{\alpha} \vdash \phi \ol{G}}
  \Eq_{\rho} [\phi := \Eq_{(T^\set_{H,\rho})^{n+1} K_0}]
  \overline{[\alpha := \Eq_A]} 
\end{split}
\]
Here, the second equality is by the induction hypothesis for
Equation~\ref{eq:iel-fix-point-intermediate2} on the $G$s, the fourth
is by the induction hypothesis for the IEL on the $G$s, and the fifth
is by Equation~\ref{eq:iel-fix-point-intermediate1}, which we have
just proved.

@@

  \Eq_{\rho} [\phi := T^{n+1}_{H,\Eq_{\rho}} K_0]

@@

&= (\mu {T_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}})

@@

  [\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]} \\ 
&= \colim{m \in \nat} {T^m_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}}\,K_0

@@

  [\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}) \\ 
&= \colim{m \in \nat} {T^m_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}}\,K_0

@@

  [\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha := \Eq_A}]}) \\  %%
&= \colim{m \in \nat} {T^m_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha := \Eq_A}]}}\,K_0

@@

  [\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha := \Eq_A}]}) \\ 
&= (\mu {T_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha := \Eq_A}]}})

@@

  [\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha := \Eq_A}]} \\ 

@@

  \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha := \Eq_{A}}]

@@

\Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha :=

@@

    \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha :=

@@

& T^{m+1}_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha := \Eq_A}]}\,K_0 \\
&= T_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha
        := \Eq_A}]} (T^{m}_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}}

@@

&= T_{G, \Eq_{\rho}[\phi := T^{n+1}_{H,\Eq_{\rho}} K_0][\overline{\alpha

@@

    \ol\beta \vdash G} \Eq_{\rho} [\phi := T^{n+1}_{H,\Eq_{\rho}}

@@

        := \Eq_{(T^{\set}_{H,\rho})^{n+1} K_0}][\overline{\alpha :=

@@

    \ol\beta \vdash G} \Eq_{\rho} [\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1}

@@

      \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1}

@@

&= T_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1}

@@

&= T^{m+1}_{G, \Eq_{\rho}[\phi := \Eq_{(T^{\set}_{H,\rho})^{n+1}

@@

      \Phi \vdash F \times G}\rho}$\qedhere

@@

@@

Expanding the second equation gives that if $x \in \setsem{\Gamma;
  \Phi \vdash F}\rho$ then
\[(\setsem{\Gamma; \Phi \vdash F}\id_{\rho}\, x, \setsem{\Gamma;
  \Phi \vdash F} f\, x) \in \relsem{\Gamma; \Phi \vdash F}\graph{f}\]
Then $\setsem{\Gamma; \Phi \vdash F}\id_{\rho}\, x =
\id_{\setsem{\Gamma; \Phi \vdash F}\rho} x = x$, so for any $x \in
\setsem{\Gamma; \Phi \vdash F}\rho$ we have $(x, \setsem{\Gamma;
  \Phi \vdash F}f\, x)$ $\in \relsem{\Gamma; \Phi \vdash F}\graph{f}$.
Moreover, $x \in \setsem{\Gamma; \Phi \vdash F}\rho$ if and only if
$(x, \setsem{\Gamma; \Phi \vdash F} f\, x) \in \graph{\setsem{\Gamma;
    \Phi \vdash F}f}$ and, if $x \in \setsem{\Gamma; \Phi \vdash
  F}\rho$ then $(x, \setsem{\Gamma; \Phi \vdash F} f\, x) \in
\relsem{\Gamma; \Phi \vdash F} \graph{f}$. Thus, if $(x, \setsem{\Gamma;
  \Phi \vdash F} f\, x) \in \graph{\setsem{\Gamma; \Phi \vdash F}f}$
then $(x, \setsem{\Gamma; \Phi \vdash F} f\, x) \in \relsem{\Gamma;
  \Phi \vdash F} \graph{f}$, i.e., $\graph{\setsem{\Gamma; \Phi \vdash
    F}f} \subseteq \relsem{\Gamma; \Phi \vdash F} \graph{f}$.  \qed

@@

    \phi.\lambda \overline \alpha.H)\overline \beta \;F)}$$ is a
  natural transformation from $\setsem{\Gamma;\emptyset \vdash
    \emptyset}$ to \[\setsem{\Gamma; \emptyset \vdash
    \phi.\lambda \overline \alpha.H)\overline \beta\,F)}\] we need
  only show that, for all $\rho : \setenv$, all $\eta :

@@

      :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F} \rho$, and the
  unique $d : \setsem{\Gamma;\emptyset \vdash \emptyset} \rho$, 

@@

by naturality of $\mathit{fold}_{T^\set_{H,\rho}}\eta$.  To see that,
for all $\rho : \setenv$, $\eta : \setsem{\Gamma; \emptyset \vdash
  \Nat^{\ol{\beta}}\,H[\phi :=_{\ol{\beta}} F][\ol{\alpha :=
      \beta}]\,F} \rho$, and $d \in \setsem{\Gamma; \emptyset \vdash
  \emptyset}\rho$,

@@

needed to be in $\setsem{\Gamma;\emptyset \vdash

@@

additional condition needed to be in $\setsem{\Gamma; \emptyset

@@

Now, since $K$ does not contain constants, $\setsem{ \emptyset;
  \ol\alpha \vdash K } (\pi_2 \rho) [\ol{\alpha := 0}] = 0$. Thus, for
any $\ol{S : \set_0}$ and $\ol{f : \setsem{ \emptyset; \ol\alpha
    \vdash K } (\pi_1 \rho)[\ol{\alpha := S}] \to \setsem{ \Gamma;
    \Phi \vdash A } (\pi_1 \rho)} $ we have relations $\ol{(S, 0, 0)}$
and morphisms $\ol{(f, !) : \relsem{ \emptyset; \ol\alpha \vdash K }
  \rho[\ol{\alpha := (S, 0, 0)}] \to \relsem{ \Gamma; \Phi \vdash A }
  \rho}$, where, for any set $X$, we suppress the sub- and
superscripts and write $!$ for the unique morphism $!^0_X$ from $0$ to
$X$. We then define
\[
k : \setsem{ \Gamma; \Phi \vdash (\Lan^{\ol \alpha}_{\ol{K}} F)\ol{A}
} (\pi_1 \rho) \to \pi_1 \,( \relsem{ \Gamma; \Phi \vdash (\Lan^{\ol
    \alpha}_{\ol{K}} F)\ol{A} } \rho )
\]
to be the unique function given by the universal property of the
colimit representation of $k$'s domain. That is, we take $k$ to be the
unique morphism such that, for any $\ol{S : \set_0}$ and \[\ol{f :
  \setsem{ \emptyset; \ol\alpha \vdash K } (\pi_1 \rho)[\ol{\alpha :=
      S}] \to \setsem{ \Gamma; \Phi \vdash A } (\pi_1 \rho)}\] we have
$k \circ j_{\ol{S},\, \ol{f}} = \pi_1 \iota_{\ol{(S, 0, 0)},\, \ol{(f,
    !)}}$.
